### 2018-12-20 writing/thesis/chapters/introduction.tex
Further work on ADER-DG, NS, AMR
Some more work.
Change stuff.

\chapter{Introduction}
We make the following contributions:

\begin{itemize}
\item We implement an \textsc{ader-dg} scheme for the compressible Navier-Stokes equations.
\item We compare different mesh refinement strategies and evaluate their computational efficiency.
\item We investigate the ability of our scheme to simulate classical fluid mechanics test scenarios correctly.
\item We describe the simulation of flows with a background state that is in hydrostatic equilibrium.
  This is useful for problems in atmospheric flow and for problems in meterology.
  Furthermore, we evaluate our discretization for these scenarios.
\end{itemize}

>>>>

\chapter{Introduction}\label{chap:introduction}
We make the following contributions:
\begin{itemize}
\item We implement an \textsc{ader-dg} and \textsc{muscl}-scheme for the compressible Navier-Stokes equations.
\item We describe a coupling of the compressible Navier-Stokes equations with an advection-diffusion-reaction term for two dimensional problems.
\item We compare different mesh refinement strategies and evaluate their computational efficiency.
\item We investigate the ability of our scheme to simulate classical fluid mechanics test scenarios correctly.
\item We describe the simulation of flows with a background state that is in hydrostatic equilibrium.
  This is useful for problems in atmospheric flow and for problems in meterology.
  Furthermore, we evaluate our discretization for these scenarios.
\end{itemize}

### 2018-12-20 writing/thesis/chapters/methods.tex
Change stuff.

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1]{\Phi_{#1}}
\newcommand{\stbasis}[1]{\Phi_{#1}}
\newcommand{\testfunction}[1]{\Phi_{#1}}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}

@@

\sidetitle{Reference Cell}
\newcommand{\mapping}{\mathcal{M}}
\newcommand{\volume}{V}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\cell = [0, 1]^D$\todo{check if $[0,1]$ detontes correct interval}.
The mapping, which takes a point on the reference cell with coordinates $\hat{\bm{x}}$ and returns a point $\bm{x}$ on a cell with center and widths $(\Delta x, \Delta y, \Delta z)$, is given by
\begin{equation}
  \hat{x} = \mapping (\bm{x}) =
  \operatorname{cell-center}(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
 % \operatorname{diag}(\Delta x, \Delta y, \Delta z)
  \begin{pmatrix}
    x - 0.5\\
    y - 0.5\\
    z - 0.5
  \end{pmatrix}.
\end{equation}
We also define the inverse mapping $\mapping^{-1}$.
This can be used to define a function $f(\bm{x})$ acting on a point $\bm{x}$ of a regular cell in term of a function $\hat{f}(\bm{x})$ that acts on a point of the reference cell by considering
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

The determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z.
\end{equation}
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }
\end{equation}

\sidetitle{Basis Functions}
\newcommand{\lagrange}[1][i]{L_{#1}}
We define basis functions on the reference cell.
We use Langrange polynomials that are defined in the one-dimensional case by
\begin{equation}
  \label{eq:lagrange-basis}
  \lagrange(x) = \prod_{i \neq j}^N \frac{x - x_j}{x_i - x_j}.
\end{equation}
We can then represent functions as a sum over the support points $x_j$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(x_i) \lagrange[i](x) = f(x_i) \lagrange[i](x),
\end{equation}
where we used the Einstein summation convention where summation over repeated indices is implied.
The support points $x_i$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(x) \dd{x} = \sum_n^N \quadWeight[n] f(x_n),
\end{equation}
where $x_n$ is a quadrature point with weight $\quadWeight[n]$.
The basis functions are similarly defined in $d$ dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature.
To simplify notation, we order the polynomials by a linearised index, a $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{N^d} f(\bm{x}_i) \lagrange[i](\bm{x}) = f(\bm{x}_i) \lagrange[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \int_0^1 f(\bm{x}) \dd{\bm{x}} &= \sum_n^{(N+1)^D - 1} \quadWeight[n] f(\bm{x}_n) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[i].
\end{align}
\todo{Fix notation, especially multi-dimensional integral.}
\todo{Check number of total polynomials.}
% TODO: Quadrature

Two important properties of this family of polynomials are
  \begin{alignat}{10}
    \label{eq:basis-interpolating}
&& \text{Interpolating:} \qquad & \lagrange[i] (x_j) &= \delta_{ij}\\
\label{eq:basis-orthogonality}
&& \text{Orthogonality:} \qquad & \ldots &= \ldots
    \end{alignat}

\sidetitle{Predictor}
To derive the predictor, we take an approximation of our solution in a nodal basis that is defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
Furthermore we assume that all quantities are given in this represantation.
We now multiply the conservation law by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \testfunction{k} \source(\stpredictor)
}}.
\end{equation}
We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells, we account for this in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \testfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \testfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
-
\intdcell{
  \testfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
} = \\
\intdt{\intdcell{
    \testfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
&+
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be by a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Collect into matrices and show scheme?}
To show the computational efficiency of our scheme, we follow the approach of~\cite{dumbser2008unified} and collect our integrals into matrices\todo{Source is Dominics \aderdg-document.}.

@@

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law-gradient} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
Note that the testfunctions
\begin{equation}
  \label{eq:cell-approx-space}
  q(\bm{x})_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}).
\end{equation}

here do not depend on time but only on the spatial dimensions.
We arrive yet again at the weak formulation of the \pde\
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(\Q, \gradQ} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(\Q, \bm{x}, t)
}}.
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with the spacetime-predictor $\stpredictor (\bm{x},t)$ and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\newcommand{\massMatrixDef}{\intdcell{
  \testfunction{k} \sbasis{l}
}}
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  \flux(\stpredictor, \gradient{\stpredictor}) % todo correct gradient?
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term is our mass matrix with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
\begin{equation}
  \label{eq:mass-matrix}
  \massMatrix[kl] = \massMatrixDef = \volume \underbrace{\quadWeight[k] \delta_{kl}}_{\text{diagonal}}
\end{equation}
which is diagonal thanks to the orthogonal basis \pcref{eq:basis-orthogonality} and thus trivial to invert.
\newcommand{\sumbasis}{\sum_i^{(N + 1)^d - 1}}
We can then insert the definition of our basis function into our equation and arrive at
\begin{align}
  \massMatrix \left( \dsol^{t+1} - \dsol^{t} \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
\begin{split}
  a &= \sumbasis \\
  b &= \sumbasis \\
  s &= \sumbasis.
\end{split}
\end{align}
We can solve this for the new unknows $\bm{u}^{n+1}$
\begin{align}\label{eq:update-predictor}
\begin{split}
  \bm{u}^{n+1} &= \bm{u}^{n} + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix^{-1} \left( \bm{a} - \bm{b} + \bm{s} \right)
\end{split}
\end{align}
which is clearly of the form of a one-step scheme.

@@

We use the total variation (\textsc{tv}) over an arbitrary indicator variable given by
\newcommand{\tv}{\operatorname{TV}}
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}] \right] =
\Vert \left(\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert } \right) \Vert_1
\end{equation}
This integral can be evaluated efficiently using \cref{eq:integration-by-substitution}.
\todo{Gradient on reference cell, Gaussian quad., inline}
\todo{Definition for FV schemes (maybe just mapping to DG-Cells)}

\newcommand{\gobs}{\operatorname{G}}
\newcommand{\mean}{\mu}
\newcommand{\std}{\sigma}
\newcommand{\variance}{\std^2}
\newcommand{\gobsCount}{n}
We want to compute the mean and the variance over all grid cells by performing only pairwise merge operations.
This allows us to minimize communication over nodes and allows for a simple implementation.
The simplest way to compute this would be by using the textbook definitions of mean and variance of a random variable $X$
\newcommand{\expectation}{\mathbb{E}}
\begin{align}
  \begin{split}
    \mu [X] &= \expectation \left[ X \right]\\
    \variance [X] &= \expectation \left[ X^2 \right] - \expectation \left[ X \right]^2.
  \end{split}
\end{align}
The reduction for the variance is numerically unstable when the variance is orders of magnitudes smaller than the mean as this can lead to catastrophic cancellations.
We thus compute the mean and variance with the parallel algorithm of~\cite{chan1982updating}.
To do this, we need to store three variables per computation unit:
the mean, the variance and the number of processed elements.
We collect these items in the vector $\gobs = (\mean, \variance, \gobsCount)$
We can then merge two pairs of observed variables (\cref{alg:merge-variance}).

>>>>

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1][]{\Phi_{#1}}
\newcommand{\stbasis}[1][]{\Theta_{#1}}
\newcommand{\sbasisRef}[1][]{\hat{\sbasis[#1]}}
\newcommand{\stbasisRef}[1][]{\hat{\stbasis[#1]}}
\newcommand{\stestfunction}[1]{\sbasis[#1]}
\newcommand{\sttestfunction}[1]{\stbasis[#1]}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}

@@

\sidetitle{Reference Cell}
\newcommand{\mapping}{\mathcal{M}}
\newcommand{\volume}{V}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\cell = [0, 1]^D$\todo{check if $[0,1]$ detontes correct interval}.
The mapping, which takes a point on the reference cell with coordinates $\hat{\bm{x}}$ and returns a point $\bm{x}$ on a cell with center and widths $(\Delta x, \Delta y, \Delta z)$, is given by
\begin{equation}\label{eq:space-mapping}
  \hat{x} = \mapping (\bm{x}) =
  \operatorname{cell-center}(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
 % \operatorname{diag}(\Delta x, \Delta y, \Delta z)
  \begin{pmatrix}
    x - 0.5\\
    y - 0.5\\
    z - 0.5
  \end{pmatrix}.
\end{equation}

We also define the inverse mapping $\mapping^{-1}$.
This can be used to define a function $f(\bm{x})$ acting on a point $\bm{x}$ of a regular cell in term of a function $\hat{f}(\bm{x})$ that acts on a point of the reference cell by considering
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

Similar to this, we define a mapping that maps a point in time to the reference time $\in (0,1)$
\todo{Fix this/remove}
\newcommand{\timeMapping}{\mathcal{T}}
\begin{equation}
  \label{eq:time-mapping}
  \hat{t} = \timeMapping(t) = t_0 + (\Delta t) t,
\end{equation}
where $\Delta t$ is timestep of a cell and $t_0$ is the beginning of its time interval.

The Jacobian determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z,
\end{equation}
the Jacobian determinant of $\timeMapping$ is $\Delta t$.
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }
\end{equation}
\sidetitle{Basis Functions}
\newcommand{\lagrangeRef}[1][i]{\hat{L}_{#1}}
\newcommand{\lagrange}[1][i]{L_{#1}}
The next step is the definition of basis functions for the reference cell.
We use Langrange polynomials that are defined in the one-dimensional case by
\begin{equation}
  \label{eq:lagrange-basis}
  \lagrangeRef(x) = \prod_{i \neq j}^N \frac{x - x_j}{x_i - x_j}.
\end{equation}
We can then represent functions as a sum over the support points $x_j$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(x_i) \lagrangeRef[i](x) = f(x_i) \lagrangeRef[i](x),
\end{equation}
where we used the Einstein summation convention where summation over repeated indices is implied.
The support points $x_i$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(x) \dd{x} = \sum_n^N \quadWeight[n] f(x_n),
\end{equation}
where $x_n$ is a quadrature point with weight $\quadWeight[n]$.
The basis functions are similarly defined in $d$ dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature.
To simplify notation, we order the polynomials by a linearised index, a $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{N^d} f(\bm{x}_i) \lagrangeRef[i](\bm{x}) = f(\bm{x}_i) \lagrangeRef[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \underbrace{\int_0^1 \cdots \int_0^1}_{D \text{ times}} f(\bm{x}) \dd{\bm{x}} &= \sum_n^{(N+1)^D - 1} \quadWeight[n] f(\bm{x}_n) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[i].
\end{align}
\todo{Fix notation, especially multi-dimensional integral.}
\todo{Check number of total polynomials.}
% TODO: Quadrature

Two important properties of this family of polynomials are
  \begin{alignat}{10}
    \label{eq:basis-interpolating}
&& \text{Interpolating:} \qquad & \lagrangeRef[i] (x_j) &= \delta_{ij}\\
\label{eq:basis-orthogonality}
&& \text{Orthogonality:} \qquad &
\int_0^1 \cdots \int_0^1 \lagrangeRef[j](\hat{\bm(x)} \lagrangeRef[j]{\hat{\bm{x}}}) \dd{\hat{\bm{x}}}  &= \quadWeight[i] \delta_{ij}
    \end{alignat}

We will later need basis functions over a space-cell 
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasisRef(\hat{\bm{x}}) &= \sum_i^{(N+1)^D - 1} \lagrange[i] (x_i)\\
  \sbasis(\bm{x}) &= \sbasisRef \left( \mapping(\bm{x}) \right)
  \end{split}
\end{align}
where $\sbasisRef$ denotes the basis function on the reference cell.
Similar we define a basis for a space-time element defined by $(0, 1)^D \times (t^n, t^{n+1})$
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef(\hat{\bm{x}}, \hat{t}) &= \sum_j^{N} \sum_i^{(N+1)^D - 1} \lagrange[j](x_j) \lagrange[i] (x_i)\\
  \stbasis (\bm{x}, t) &= \stbasisRef \left( \mapping(\bm{x}), \timeMapping(t) \right)
  \end{split}
\end{align}

\sidetitle{Predictor}
To derive the predictor, we take an approximation of our solution in a nodal basis that is defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
Furthermore we assume that all quantities are given in this represantation.
We now multiply the conservation law by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \sttestfunction{k} \source(\stpredictor)
}}.
\end{equation}
We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells, we account for this in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
-
\intdcell{
  \sttestfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
} = \\
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
&+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be by a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Collect into matrices and show scheme?}
To show the computational efficiency of our scheme, we follow the approach of~\cite{dumbser2008unified} and collect our integrals into matrices\todo{Source is Dominics \aderdg-document.}.

@@

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law-gradient} by a test function $\stestfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
Note that the testfunctions
\begin{equation}
  \label{eq:cell-approx-space}
  q(\bm{x})_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}).
\end{equation}

here do not depend on time but only on the spatial dimensions.
We arrive yet again at the weak formulation of the \pde\
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction{k} \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction{k} S(\Q, \bm{x}, t)
}}.
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with the spacetime-predictor $\stpredictor (\bm{x},t)$ and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction{k} \sbasis[l]
}}
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \stestfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\stestfunction{k}} \cdot  \flux(\stpredictor, \gradient{\stpredictor}) % todo correct gradient?
}}\right)
&=
\left(\intdt{\intdcell{
      \stestfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term is our mass matrix with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
\begin{equation}
  \label{eq:mass-matrix}
  \massMatrix[kl] = \massMatrixDef = \volume \underbrace{\quadWeight[k] \delta_{kl}}_{\text{diagonal}}
\end{equation}
which is diagonal thanks to the orthogonal basis \pcref{eq:basis-orthogonality} and thus trivial to invert.
\newcommand{\sumbasis}{\sum_i^{(N + 1)^d - 1}}
We can then insert the definition of our basis function into our equation and arrive at
\begin{align}
  \massMatrix \left( \dsol^{t+1} - \dsol^{t} \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
\begin{split}
  a &= \sumbasis F_n \intdcell{
\sbasis[n] \gradient{\sbasis[n]}
  } \\
  b &= \ldots \\
  s_n &= \sumbasis \intdcell{
s \sbasis[n]
  }.
\end{split}
\end{align}
We can solve this for the new unknows $\bm{u}^{n+1}$
\begin{align}\label{eq:update-predictor}
\begin{split}
  \bm{u}^{n+1} &= \bm{u}^{n} + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix^{-1} \left( \bm{a} - \bm{b} + \bm{s} \right)
\end{split}
\end{align}
which is clearly of the form of a one-step scheme.

@@

Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a function that maps the discrete solution of a cell to an arbitrary indicator variable.
Furthermore, we assume that $\bm{f}$ is expanded in the space-basis as in \cref{eq:cell-approx-space}.
Then the total variation (\textsc{tv}) of this function is defined by
\newcommand{\tv}{\operatorname{TV}}
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}] \right] =
\left(\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert } \right).
\end{equation}
This integral can be evaluated efficiently on the reference cell \pcref{eq:space-mapping,eq:cell-approx-space} with Gaussian quadrature \pcref{eq:quadrature}.

\todo{Gradient on reference cell, Gaussian quad., inline}
\todo{Definition for FV schemes (maybe just mapping to DG-Cells)}
\newcommand{\gobs}{\operatorname{G}}
\newcommand{\mean}{\mu}
\newcommand{\std}{\sigma}
\newcommand{\variance}{\std^2}
\newcommand{\gobsCount}{n}
We want to compute the mean and the variance over all grid cells by performing only pairwise merge operations.
This allows us to minimize communication over nodes and allows for a simple implementation.
The simplest way to compute this would be by using the textbook definitions of mean and variance of a random variable $X$
\newcommand{\expectation}{\mathbb{E}}
\begin{align}
  \begin{split}
    \mu [X] &= \expectation \left[ X \right]\\
    \variance [X] &= \expectation \left[ X^2 \right] - \expectation \left[ X \right]^2.
  \end{split}
\end{align}
The reduction for the variance is numerically unstable when the variance is orders of magnitudes smaller than the mean as this can lead to catastrophic cancellations.
We thus compute the mean and variance with the parallel algorithm of~\cite{chan1982updating}.
To do this, we need to store three variables per computation unit:
the mean, the variance and the number of processed elements.
We collect these items in the vector $\gobs = (\mean, \variance, \gobsCount)$
We can then merge two pairs of observed variables (\cref{alg:merge-variance}).
